---
title: "Lab1"
author: "Colter King"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tasks

## Task 1

Find the working directory

```{r}
getwd()
```

## Task 2

Read in the DDT data

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```

## Task 3

  a. The qualitative variables are RIVER and SPECIES.
  
  b. The quantitative variables are MILE, LENGTH, WEIGHT, and DDT.
  
  c. There are three species in the data set.
  
  d. Subset the data so only Lumbass over 800 grams are accounted for.
  
```{r}
LUMBASS800=with(ddt, ddt[WEIGHT>800 & SPECIES == "LMBASS",])
summary(LUMBASS800)
```

  e.  Subset the data so only fish in the SCM river with DDT 4.0 and higher are accounted for.
  
```{r}
SMC4 = with(ddt, ddt[DDT >4.0 & RIVER == "SCM",])
summary(SMC4)
```

## Clicker Questions

```{r}
mean(ddt$LENGTH)
``` 

  a. The mean length is 42.81.
  
```{r}
sd(ddt$WEIGHT)
``` 

  b. The standard deviation of the fish weights is 376.55.
  
  c. No,The values are correlated with the wrong variables as displayed.
  
  d. The last value for v/20 would be 1.0
```{r}
  v=1:20
  v/20
```


## Task 4

  A table of the rivers followed by a barplot of the same data.
  
```{r}
sp=with(ddt,table(RIVER))
sp
barplot(sp,beside=TRUE,col=1:4)
```

  A table of the rivers crossed with the fish species followed by a bar graph of the same data.
  
```{r}
spish=with(ddt,table(RIVER,SPECIES))
spish
barplot(spish,beside=TRUE,col=5:8)
```

## Task 5

  A pie chart displaying the Species of fish.

```{r}
spec=with(ddt,table(SPECIES))
pie(spec, col = 9:11)
```

  A pie chart of the Rivers fish were found.

```{r}
riv=with(ddt,table(RIVER))
pie(riv, col = 12:15)
```


## Task 6

  Box plot of the DDT data.

```{r}
boxplot(ddt$DDT, col="Green", main="DDT")
```
  
  Box plot of the fish Weight.

```{r}
boxplot(ddt$WEIGHT, col="RED", main="Weight")
```

  Box plot of the fish Length.

```{r}
boxplot(ddt$LENGTH, col="BLUE", main="Length")
```

## Task 7

  Co-plot comparing the Length vs. Weight of the fish by their River.

```{r}
with(ddt, coplot(LENGTH~WEIGHT|RIVER))
```
  
  Co-plot comparing the DDT vs. Weight of the fish by their Species.

```{r}
with(ddt, coplot(DDT~WEIGHT|SPECIES))
```


## Task 8

  Box plot of the Species vs. Weight catagorized by their River.

```{r}
library(ggplot2)
g = ggplot(ddt, aes(x=SPECIES , y=WEIGHT, fill=RIVER))
g = g + geom_boxplot() + ggtitle("Colter King")
g
```

  Violin plot of the Species vs. Weight catagorized by their River.

```{r}
h = ggplot(ddt, aes(x=RIVER , y=LENGTH, fill=SPECIES))
h = h + geom_violin() + ggtitle("Colter King")
h
```

  Point plot of the Weight vs. Length of fish catagorized by their species.

```{r}
h = ggplot(ddt, aes(x=WEIGHT , y=LENGTH, fill=SPECIES, color = SPECIES))
h = h + geom_point() + ggtitle("Colter King")
h
```
