---
title: "Lab2"
author: "Colter King"
date: "9/3/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tasks



## Task 1

Find the working directory

```{r}
getwd()
```

## Task 2

Read in EPAGAS deta and display the first 6 points.

```{r}
EPAGAS = read.csv("EPAGAS.csv")
head(EPAGAS)
```

## Task 3

Makes the object mpg

```{r}
mpg = EPAGAS$MPG
head(mpg)
```

Stores z values for elements of mpg data

```{r}
z = ( (mpg - mean(mpg)) / sd(mpg) )
head(z)
```

Shows the average of the z values of the data set is zero and their variance is 1.

```{r}
print(paste0("mean = ", round(mean(z), digits = 4)), quote = FALSE)
```

```{r}
print(paste0("variance = ", var(z)), quote = FALSE)
```

The following mpg values are possible outliers to the data set

```{r}
mpg[abs(z) >= 2 & abs(z) <= 3]
```

The following mpg values are defined as outliers

```{r}
mpg[abs(z) >= 3]
```



```{r}
mpgcolor = ifelse(abs(z)>3, "Red", 
               ifelse(abs(z)>=2 & abs(z)<=3, "Blue", 
                      "Black")) 

library(lattice)
dotplot(mpg, col = mpgcolor, xlab = "MPG", main = "MPG values")
```

## Task 4

```{r}
boxplot(mpg, main = "MPG Boxplot", col = "yellow2", xlab = "MPG", horizontal = TRUE, notch = TRUE)
```

### Chebyshev's Theorem
Chebyshev's themoem represented by
$$1 - \frac{1}{k^2} $$
predicts what percentage of our data will fall within k standard deviations of the mean. Below we use the theorem 

Chebyshev's predicted proportion of data within two standard deviations:

```{r}
k = 2
chebyshev = 1 - 1/(k^2)
chebyshev
```


The exact proportion of data within  standard deviations:
```{r}
extprop = length(mpg[abs(z) < 2]) / length(mpg)
extprop
```

### Empirical Rule

The empirical rule states that in any normal distribution nearly all data will fit within three standard deviations of the mean. It states that approximately 95% of the data will fall within two standard deviations if it is a normal distribution. 

```{r}
empirical = length(mpg[mean(mpg)-2*sd(mpg) < mpg & mpg < mean(mpg)+ 2*sd(mpg) ])/length(mpg)
empirical
```
Based on the calculations above, the empirical rule is valid as 96% of the data falls within the two standard deviations the mean which is slightly greater than the predicted 95% the empirical rule defines. This means the data set follows a normal distribution which can be visualized by the boxplot in section 4 and the dot plot in section 3.









